UCD\\MSc Thesis Proposal:\\\emph{An approach for Continuous Capacity Planning in Cloud Environments with Uptime-based Pricing Model}

### 2015-04-22 paper/paper.tex
updates

\myparagraph{Visual recognition}

Visual recognition as a helper method to define thresholds, especially for further incremental optimizations. 
Useful as static thresholds can miss a whole set of machines that are just below the threshold.

\begin{figure}[H]
       \includegraphics[width=\linewidth]{figures/cubism}
      \caption{Visualization of distance from optimality points. Green - positive, Blue - negative values.}
        \label{fig:cubism}
\end{figure}

@@

Settings were tweaked during computation mostly due to frequent Out Of Memory errors. Default driver and executor settings - 256M and 512M respectively - were not enough to handle the volume of input data. Further OOMs occurred during \textit{RRD.reduceByKey}\footnote{\url{https://spark.apache.org/docs/1.2.1/api/python/pyspark.html#pyspark.RDD.reduceByKey}} operations as it required to compute a fairly large hash map. Tweaks of \textit{spark.default.parallelism} setting worked the best as it enforced more granular input data chunking.

@@

The majority of machines within Google cluster i.e. 82\%, have following configuration: 0.5 CPU, 0.5 RAM or 0.5 CPU, 0.25 RAM. We will assign heavy pricing scheme for these machines. For the rest of the cluster, we will assign medium pricing scheme.

@@

Transformation to below schema. 
- total number of tasks per day

@@

Key distance indicator got calculated for both schemes based on Algorithm~\ref{alg:distance_from_optimality}.

\begin{table}[h]
\begin{center}
    \begin{tabular}{| l | l | l |}
    \hline
    Indicator & Scheme 1 & Scheme 2 \\
    \hline
    Total cost & \$8 177 048 & \$5 637 418 \\
    \hline
    KDI & +3 215 231 & +5 431 137 \\
    \hline
%    Total CPU & 15972217.38 & 15972217.38 \\
%    \hline
%    Total Memory & 30676363.36 & 30676363.36 \\
%    \hline
%    Number of tasks & 1234716525 & 1234716525 \\
%    \hline
    \end{tabular}
\end{center}
\caption{KDI and total cost for cluster}
\label{tab:kdi_and_cost}
\end{table}

\myparagraph{Cluster SLO averages}

Avg cpu usage per machine = 15972217.38/12583 = 1269.35/29 = 43.77
Avg mem usage per machine = 30676363.36/12583 = 2437.92/29 = 84.06
Avg tasks number per machine = 1234716525/12583 = 98125.77/29 = 3383.65

\section{Recognize inefficiencies with Threshold-based Report}

The report will filter all machines that reported non-optimal - negative distance from intersection point - configuration during measurement period (29 days).

\myparagraph{Scheme 1}
Number of machines to tune: 932
CPU capacity to accommodate: 1 320 781.36
Memory capacity to accommodate: 4 286 742.06
Number of tasks to execute: 118869614
Cost of running non-optimal machines: \$657 200
Total uptime: 648519
cost per hour = \$1.01338588383687
Cost per task =  \$657 200 / 118869614 = \$0.00552874681666

\myparagraph{Scheme 2}
Number of machines to tune: 937
CPU capacity to accommodate: 1 327 202.18
Memory capacity to accommodate: 4 303 983.31
Number of tasks to execute: 119367988
Cost of running non-optimal machines: \$532 840.76
Total uptime: 651899

@@

@@

vm1 = Entity(c_name, c_light, c_medium, c_heavy)
vm2 = Entity(m_name, m_light, m_medium, m_heavy)

>>> calc_options(uptime, vm2, 29)
(1597, 3727, 972, 567 273.49, 998 186.23, 410 041.68)
>>> calc_options(uptime, vm1, 29)
(1397, 1720, 972, 770 417.58, 876 376.34, 604 274.61)

>>>>

@@

Settings were tweaked during computation mostly due to frequent Out Of Memory errors. Default driver and executor settings - 256M and 512M respectively - were not enough to handle the volume of input data. Further OOMs occurred during \textit{RRD.reduceByKey}\footnote{\url{https://spark.apache.org/docs/1.2.1/api/python/pyspark.html\#pyspark.RDD.reduceByKey}} operations as it required to compute a fairly large hash map. Tweaks of \textit{spark.default.parallelism} setting worked the best as it enforced more granular input data chunking.

@@

The majority of machines within Google cluster i.e. 82\%, have following configuration: 0.5 CPU, 0.5 RAM or 0.5 CPU, 0.25 RAM. We will assign heavy pricing scheme for these machines. For the rest of the cluster, we will assign medium pricing scheme. 
As a result of the assignment cluster machines are grouped into 10595 machines of heavy and 2258 of medium usage.

@@

@@

Key distance indicator was calculated based on Algorithm~\ref{alg:distance_from_optimality}. 
Total cost was calculated based on total uptime and pricing scheme assignment. Results are presented on Table~\ref{tab:kdi_and_cost}.

\begin{table}[h]
\begin{center}
    \begin{tabular}{| l | l | l |}
    \hline
    \textbf{Indicator} & \textbf{Scheme 1} & \textbf{Scheme 2} \\
    \hline
    Total cost & \$8 177 048 & \$5 637 418 \\
    \hline
    KDI & +3 215 231 & +5 431 137 \\
    \hline
%    Total CPU & 15972217.38 & 15972217.38 \\
%    \hline
%    Total Memory & 30676363.36 & 30676363.36 \\
%    \hline
%    Number of tasks & 1234716525 & 1234716525 \\
%    \hline
    \end{tabular}
\end{center}
\caption{KDI and total cost for cluster}
\label{tab:kdi_and_cost}
\end{table}

KDI for Scheme 1 and 2 is a positive number, which is expected based on 82\% of heavy scheme assignment and 23.36 hours of uptime per day on average.
Scheme 1 is 31\% more expensive than Scheme 2 and this is expected as well as \textit{cph} values are higher for Scheme 1.

\section{Recognize inefficiencies using threshold-based report}

KDI can be used to find inefficiencies.  Any value below 0 indicates non-optimal usage for a given scheme. 
A user can specify a threshold to find non-optimal cases. We will assume that non-optimal usage within the cluster is when machine reports negative KDI values for the whole trace period.
 These criteria identified 932 and 937 machines for Scheme 1 and 2 respectively. Detailed results splitted by cost and resource utilization are presented in Table~\ref{tab:threshold_based_report}.

\begin{table}[h]
\begin{center}
    \begin{tabular}{| l | l | l |}
    \hline
    \textbf{Property} & \textbf{Scheme 1} & \textbf{Scheme 2} \\
    \hline
    Number of machines to tune & 932 & 937 \\
    \hline
    Total cost & \$657 200 & \$532 840.76 \\
    \hline
    CPU capacity to accommodate & 1 320 781.36 & 1 327 202.18 \\
    \hline
    Memory capacity to accommodate & 4 286 742.06 & 4 303 983.31 \\
    \hline
    Number of tasks to execute & 118869614 & 119367988 \\
    \hline
    Total uptime & 648519 & 651899 \\
    \hline
    Average cost per hour & \$1.0134 & \$0.817 \\
    \hline
    Average cost per task & \$0.00553  & \$0.00446 \\
    \hline
    \end{tabular}
\end{center}
\caption{Results of report run on cluster with criteria of KDI below 0 for 29 days}
\label{tab:threshold_based_report}
\end{table}

The cost of running such identified non-optimal machines is around 8\% and 10\% of total cost for Scheme 1 and 2 respectively.

@@

Once inefficiencies are identified, we will attempt to find a more optimal solution.

@@

$vm1 = Entity(c_name, c_light, c_medium, c_heavy)$
$vm2 = Entity(m_name, m_light, m_medium, m_heavy)$

$calc_options(uptime, vm2, 29)$
(1597, 3727, 972, 567 273.49, 998 186.23, 410 041.68)
$calc_options(uptime, vm1, 29)$
(1397, 1720, 972, 770 417.58, 876 376.34, 604 274.61)
